Distributional Trees for Circular Data

### 2019-02-07 pkg/disttree/inst/IWSM2019/abstract_IWSM2019_schlosser.tex
extended information on authors' intstitutes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2449 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: cc4c8e6115cdbcae47a3b44112f52319eff46c2e

% Authors and running list of authors to be used as right header:
\author{Lisa Schlosser\inst{1}, 
Moritz N. Lang\inst{1},
Torsten Hothorn\inst{2},
Georg J. Mayr\inst{1},
Reto Stauffer\inst{1},
Achim Zeileis\inst{1}}
\authorrunning{Schlosser et al.}    %% use \authorrunning{Surname 1} if only 1 author
                                    %% use \authorrunning{Surname 1 and Surname2} if two authors
                                    %% use \authorrunning{Surname 1 et al.} if more than two authors

% Institutes of all authors
% Include city and country of each institute, do not include the full address.
\institute{Universit\"at Innsbruck, Austria
\and Universit\"at Z\"urich, Switzerland}

@@

\begin{figure}[p!]\centering
\includegraphics[height = .5\textheight,angle=90,origin=c]{schlosser-circtree_plot.pdf}
\caption{Fitted tree based on the von Mises distribution for wind direction forecasting.
In each terminal node the empirical histogram (gray) and fitted density (red line)
are depicted along with the estimated location parameter (red hand). The covariates
employed are wind direction ($\text{rad}$), wind speed ($\text{ms}^{-1}$),
and pressure gradients ($\text{hPa}$) west and east of the airport,
all lagged by one hour.}
\label{schlosser:fig_tree} \end{figure}

@@

Figure~\ref{schlosser:fig_tree} depicts the resulting distributional tree, including
both the empirical (gray) and fitted von Mises (red) distribution of wind direction
in each terminal node. Based on the fitted location parameters $\hat \mu$, the subgroups
can be distinguished into the following wind regimes:
(1)~Up-valley winds blowing from the valley mouth towards the upper valley
(from east to west, nodes 4 and 5). (2) Downslope winds blowing across
the Alpine crest along the Wipp Valley towards Innsbruck (from south-east to
north-west, nodes 7 and 8). (3) Down-valley winds blowing in the direction
of the valley mouth (from west to east, nodes 12, 14, and 15). Node~11
captures observations with rather low wind speeds that cannot be distinguished
clearly into wind regimes and consequently are associated with a very low
estimated concentration $\hat \kappa$. In terms of covariates, the lagged
wind direction (``persistence'') is mostly responsible for distinguishing
the broad wind regimes listed above while the pressure gradients and wind
speed separate between subgroups with high vs.\ low precision.

@@

>>>>

% Authors and running list of authors to be used as right header:
\author{Lisa Schlosser\inst{1}, 
Moritz N. Lang\inst{{1,2}},
Torsten Hothorn\inst{3},
Georg~J.~Mayr\inst{2},
Reto Stauffer\inst{1},
Achim Zeileis\inst{1}}
\authorrunning{Schlosser et al.}    %% use \authorrunning{Surname 1} if only 1 author
                                    %% use \authorrunning{Surname 1 and Surname2} if two authors
                                    %% use \authorrunning{Surname 1 et al.} if more than two authors

%% Institutes of all authors
%% Include city and country of each institute, do not include the full address.
%\institute{Universit\"at Innsbruck, Austria
%\and Universit\"at Z\"urich, Switzerland}

%%%%%%%% IF DEPARTMENTS ARE LISTED AS WELL CHANGE NUMBERING!!!  %%%%%%%%%%%%%%
\institute{Department of Statistics, Universit\"at Innsbruck, Innsbruck, Austria
\and Department of Atmospheric and Cryospheric Science, Universit\"at Innsbruck, Innsbruck, Austria
\and Epidemiology, Biostatistics and Prevention Institute, Universit\"at Z\"urich, Zurich, Switzerland}

@@

\begin{figure}[p!]\centering
\includegraphics[height = .5\textheight,angle=90,origin=c]{schlosser-circtree_plot.pdf}
\caption{Fitted tree based on the von Mises distribution for wind direction forecasting.
In each terminal node the empirical histogram (gray) and fitted density (red line)
are depicted along with the estimated location parameter (red hand). The covariates
employed are wind direction ($degree$), wind speed ($\text{ms}^{-1}$),
and pressure gradients ($\text{hPa}$) west and east of the airport,
all lagged by one hour.}
\label{schlosser:fig_tree} \end{figure}

@@

Figure~\ref{schlosser:fig_tree} depicts the resulting distributional tree, including
both the empirical (gray) and fitted von Mises (red) distribution of wind direction
in each terminal node. Based on the fitted location parameters $\hat \mu$, the subgroups
can be distinguished into the following wind regimes:
(1)~Up-valley winds blowing from the valley mouth towards the upper valley
(from east to west, nodes 4 and 5). (2) Downslope winds blowing across
the Alpine crest along the intersecting valley towards Innsbruck (from south-east to
north-west, nodes 7 and 8). (3) Down-valley winds blowing in the direction
of the valley mouth (from west to east, nodes 12, 14, and 15). Node~11
captures observations with rather low wind speeds that cannot be distinguished
clearly into wind regimes and consequently are associated with a very low
estimated concentration $\hat \kappa$. In terms of covariates, the lagged
wind direction (``persistence'') is mostly responsible for distinguishing
the broad wind regimes listed above while the pressure gradients and wind
speed separate between subgroups with high vs.\ low precision.

@@

% Acknowledgments, if needed:
\acknowledgments{Special Thanks to ... }
